The \dealii Library, Version 9.1

### 2019-05-16 9.1/paper.tex
polish Overview, add some unfinished sections

The major changes of this release are:
\begin{itemize}
  \item Improved support for automatic and symbolic differentiation (see
    Section~\ref{subsec:ad}),
  \item Full support for $hp$ adaptivity in parallel computations (see
    Section~\ref{subsec:hp}),
  \item Interfaces to the HDF5 file format and libraries (see
    Section~\ref{subsec:hdf5}),
  \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
  \item Parallel geometric multigrid (GMG) improvements (see
    \cite{ClevengerHeisterKanschatKronbichler2019} and
    Section~\ref{subsec:gmg}),
  \item Four new tutorial programs (step-61, step-62, step-63, step-64),
    as well as one new code gallery program (see
    Section~\ref{subsec:steps}).
\end{itemize}
These will all be discussed in more detail in the
following section. There are a number of other noteworthy changes in this release:
\begin{itemize}
\item A number of bug fixes and performance improvements to the
  matrix-free framework and related geometric multigrid solvers. In
  particular, the implementation of the Chebyshev iteration, an often
  used smoother in the matrix-free context, has been revised to reduce
  vector accesses. This speeds up matrix-free
  multigrid solvers by up to 10-15\% on geometries with affine
  (parallelogram and parallelpiped) cells, and up to
  5\% on geometries with cells bounded by curved edges and faces.
\item Various variants of geometric multigrid solvers and matrix-free
  implementations were run on up to 304,128 cores during the acceptance phase of
  the SuperMUC-NG supercomputer, verifying the scalability of our
  implementations. Some performance bottlenecks on distributed triangulation
  with more than 100k MPI ranks have been fixed.
\marginpar{Luca: Please expand}
\item ParsedConvergenceTable.
\item The \texttt{FE\_BernardiRaugel} class implements the
  non-standard Bernardi-Raugel (BR) element that can be used as one part of
  a stable velocity/pressure pair for the Stokes equation
  \cite{BR85}. The BR element can be seen as either an enriched
  version of the $Q_1^d$ element with added bubble functions on each
  edge (in 2d) or face (in 3d), or as a reduced version of the $Q_2^d$
  element. It addresses the fact that the $Q_1^d\times Q_0$
  combination is not inf-sup stable (requiring a larger velocity
  space), and that the $Q_2^d\times Q_1$ combination is stable but
  sub-optimal since the velocity space is too large relative to the
  pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

  The current implementation of the element only works for
  parallelogram meshes due to difficulties associated with the mapping
  of shape functions: The shape functions of the $Q_1^d$ part of the
  element need to be mapped as scalars, as is common for the vector
  components of the $Q_1^d$ element; on the other hand, the
  vector-valued edge bubble functions need to be mapped using the
  Piola transform as is common for the Raviart-Thomas
  element. \dealii{} does not currently have the ability to use
  different mappings for individual shape functions, though this
  functionality is planned for the next release.

\item The \texttt{FE\_NedelecSZ} class is an alternative
  implementation of the ideas behind the N{\'e}d{\'e}elec element. It is based
  on the work of Zaglmayr \cite{Zag06} and overcomes the
  sign conflict issues present in traditional N{\'e}d{\'e}lec
  elements that arise from the edge and face parameterizations used
  in the basis functions. Therefore, this element should provide
  consistent results for general quadrilateral and hexahedral
  elements for which the relative orientations of edges and faces as
  seen from all adjacent cells are often difficult to establish.
  The way this element addresses the sign conflict problem is to assign
  local edges and faces a globally defined orientation. The
  implementation is described in \cite{Kynch2017}.

  \marginpar{Maybe also mention the rtree interfaces?}
\end{itemize}

Beyond these changes, the changelog lists more than 200 other features and bugfixes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Significant changes to the library}

This release of \dealii contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}{
in the file that lists all changes for this release}, see \cite{changes91}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

>>>>

The major changes of this release are:
%
\begin{itemize}
  \item Improved support for automatic and symbolic differentiation (see
    Section~\ref{subsec:ad}),
  \item Full support for $hp$ adaptivity in parallel computations (see
    Section~\ref{subsec:hp}),
  \item Interfaces to the HDF5 file format and libraries (see
    Section~\ref{subsec:hdf5}),
  \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
  \item Parallel geometric multigrid (GMG) improvements (see
    \cite{ClevengerHeisterKanschatKronbichler2019} and
    Section~\ref{subsec:gmg}),
  \item Four new tutorial programs (step-61, step-62, step-63, step-64),
    as well as one new code gallery program (see
    Section~\ref{subsec:steps}).
\end{itemize}
%
The major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item
  The release contains bug fixes and performance improvements of the
  matrix-free framework and related geometric multigrid solvers. In
  particular, the implementation of the Chebyshev iteration, an often used
  smoother in the matrix-free context, has been revised to reduce vector
  accesses. This speeds up matrix-free multigrid solvers by up to 10-15\%
  on geometries with affine (parallelogram and parallelpiped) cells, and up
  to 5\% on geometries with cells bounded by curved edges and faces.

\item
  Various variants of geometric multigrid solvers and matrix-free
  implementations were run on up to 304,128 cores during the acceptance
  phase of the SuperMUC-NG supercomputer, verifying the scalability of our
  implementations. Some performance bottlenecks on distributed
  triangulation with more than 100k MPI ranks have been fixed.

\item
  A new class \texttt{ParsedConvergenceTable} has been introduced.
  \textcolor{red}{TODO}
  \marginpar{Luca: Please expand}

\item
  The \texttt{FE\_BernardiRaugel} class implements the non-standard
  Bernardi-Raugel (BR) element that can be used to construct a stable
  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
  element is an enriched version of the $Q_1^d$ element with added bubble
  functions on each edge (in 2d) or face (in 3d). It addresses the fact
  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
  larger velocity space), and that the $Q_2^d\times Q_1$ combination is
  stable but sub-optimal since the velocity space is too large relative to
  the pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

  The element is currently only implemented for parallelogram meshes due to
  difficulties associated with the mapping of shape functions: The shape
  functions of the $Q_1^d$ part of the element need to be mapped as
  scalars, as is common for the vector components of the $Q_1^d$ element;
  on the other hand, the vector-valued edge bubble functions need to be
  mapped using the Piola transform as is common for the Raviart-Thomas
  element. \dealii{} does not currently have the ability to use different
  mappings for individual shape functions, though this functionality is
  planned for the next release.

\item
  The \texttt{FE\_NedelecSZ} class is a new implementation of the
  N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
  the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
  present in traditional N{\'e}d{\'e}lec elements that arise from the edge
  and face parameterizations used in the basis functions. Therefore, this
  element should provide consistent results for general quadrilateral and
  hexahedral elements for which the relative orientations of edges and
  faces (as seen from all adjacent cells) are often difficult to establish.
  The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
  assigning a globally defined orientation to local edges and faces. A
  detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
  element in \dealii{} can be found in \cite{Kynch2017}.

\item
  A new interface to \texttt{boost::geometry::index::rtree} has been added.
  \textcolor{red}{TODO}
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Major changes to the library}
\label{sec:major}

This release of \dealii contains a number of large and significant changes
that will be discussed in this section.

It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}{
in the file that lists all changes for this release}, see \cite{changes91}.
(The file is also linked to from the web site of each release as well as
the release announcement.)
